\section{Direct and Semidirect Products and Abelian Groups}

\subsection{Direct Products}

\begin{exercise}\label{ex5.1.1}
    Show that the center of a direct product is the direct product of the centers:
    \[Z(G_1 \times G_2 \times \cdots \times G_n) = Z(G_1) \times Z(G_2) \times \cdots \times Z(G_n)\]
    Deduce that the direct product of two groups is abelian if and only if each of the factors is abelian.
\end{exercise}

\begin{sol}
    Let $G = G_1 \times G_2 \times \cdots \times G_n$. Let $z = (z_1, z_2, \ldots, z_n) \in Z(G)$. Then for any $g = (g_1, g_2, \ldots, g_n) \in G$, we have $zg = gz$, or equivalently,
    \[(z_1g_1, z_2g_2, \ldots, z_ng_n) = (g_1z_1, g_2z_2, \ldots, g_nz_n)\]
    This implies that $z_i g_i = g_i z_i$ for all $i$, so $z_i \in Z(G_i)$ for all $i$. Thus, $z \in Z(G_1) \times Z(G_2) \times \cdots \times Z(G_n)$, and we have shown that $Z(G) \subseteq Z(G_1) \times Z(G_2) \times \cdots \times Z(G_n)$.

    Conversely, let $z = (z_1, z_2, \ldots, z_n) \in Z(G_1) \times Z(G_2) \times \cdots \times Z(G_n)$. Then for any $g = (g_1, g_2, \ldots, g_n) \in G$, we have
    \[zg = (z_1g_1, z_2g_2, \ldots, z_ng_n) = (g_1z_1, g_2z_2, \ldots, g_nz_n) = gz\]
    Thus, $z \in Z(G)$, and we have shown that $Z(G_1) \times Z(G_2) \times \cdots \times Z(G_n) \subseteq Z(G)$.
\end{sol}

\begin{exercise}\label{ex5.1.2}
    Let $G_1, G_2, \ldots, G_n$ be groups and $G = G_1 \times \cdots \times G_n$. Let $I$ be a proper, nonempty subset of $\set{1, \ldots, n}$ and let $J = \set{1, \ldots, n} - I$. Define $G_I$ to be the set of elements of $G$ that have the identity $G_j$ in position $j$ for all $j \in J$.
    \begin{subproblems}
        \item Prove that $G_I$ is isomorphic to the direct product of the $G_i$ for $i \in I$.
        \item Prove that $G_I$ is a normal subgroup of $G$ and that $G/G_I \cong G_J$.
        \item Prove that $G \cong G_I \times G_J$.
    \end{subproblems}
\end{exercise}

\begin{solalph}
    \item Identifying $G_i$ with the subgroup of $G$ defined in Proposition 5.2, let $H$ be the product of the $G_i$ for $i \in I$ in $G$. Let $\phi : G_I \to H$ be the map defined by mapping the non-identity coordinates of $G_I$ to the corresponding coordinates of $H$. It is clear that $\phi$ is a homomorphism. Moreover, $\phi$ is surjective by definition of $H$. Finally, the kernel of $\phi$ is the set of all elements of $G_I$ that have the identity in all coordinates, which is just the identity element of $G$. Thus, $\phi$ is an isomorphism from $G_I$ to $H$, and we are done.
    \item Consider the map $\phi : G \to G_J$ defined by mapping the non-identity coordinates of $G$ to the corresponding coordinates of $G_J$. It is clear that $\phi$ is a homomorphism. Moreover, $\phi$ is surjective by definition of $G_J$. Finally, the kernel of $\phi$ is the set of all elements of $G$ that have the identity in all coordinates corresponding to $J$, which is exactly $G_I$. Thus, $\phi$ induces an isomorphism from $G/G_I$ to $G_J$, and we are done.
    \item Any element of $G_I \times G_J$ can be written as the product of an element of $G_I$ and an element of $G_J$ by taking the first and second coordinates, respectively. Thus, the map $\phi : G_I \times G_J \to G$ defined by $\phi((g_i)_{i \in I}, (g_j)_{j \in J}) = (g_1, g_2, \ldots, g_n)$ is a bijection. Injectivity is clear, since $\ker\phi$ is trivial. It is also a homomorphism since the group operation in $G$ is defined coordinate-wise. Thus, $\phi$ is an isomorphism from $G_I \times G_J$ to $G$, and we are done.
\end{solalph}

\np

\begin{exercise}\label{ex5.1.3}
    Under the notation of the preceding exercise, let $I$ and $K$ be any disjoint nonempty subsets of $\set{1, 2, \ldots, n}$, and let $G_I$ and $G_K$ be the subgroups of $G$ defined above. Prove that $xy = yx$ for all $x \in G_I$ and $y \in G_K$.
\end{exercise}

\begin{sol}
    Let $x \in G_I$ and $y \in G_K$. Then $x$ has the identity in all coordinates corresponding to $J = \set{1, 2, \ldots, n} - I$, and $y$ has the identity in all coordinates corresponding to $L = \set{1, 2, \ldots, n} - K$. Since $I$ and $K$ are disjoint, we have $J \cup L = \set{1, 2, \ldots, n}$, so every coordinate of $x$ and $y$ is the identity in at least one of them. Thus, when we compute the product $xy$, we have
    \[xy = (x_1y_1, x_2y_2, \ldots, x_ny_n) = (y_1x_1, y_2x_2, \ldots, y_nx_n) = yx\]
    as desired.
\end{sol}

\begin{exercise}\label{ex5.1.4}
    Let $A$ and $B$ be finite groups, and let $p$ be a prime. Prove that any Sylow $p$-subgroup of $A \times B$ is of the form $P \times Q$, where $P \in \syl_p(A)$ and $Q \in \syl_p(B)$. Prove that $n_p(A \times B) = n_p(A)n_p(B)$. Generalize both of these results to a direct product of any finite number of finite groups (so that the number of Sylow $p$-subgroups of a direct product is the product of the numbers of Sylow $p$-subgroups of the factors).
\end{exercise}

\begin{sol}
    Let $|A| = p^\alpha m$ and $|B| = p^\beta n$, where $p \nmid m$ and $p \nmid n$. Then $|A \times B| = |A||B| = p^{\alpha + \beta}mn$. Let $R \in \syl_p(A \times B)$. Then $|R| = p^{\alpha + \beta}$, so $R$ is a Sylow $p$-subgroup of $A \times B$. Let $\pi_A : A \times B \to A$ and $\pi_B : A \times B \to B$ be the projection maps. Then $\pi_A(R)$ is a $p$-subgroup of $A$, so $\pi_A(R) \subseteq P$ for some $P \in \syl_p(A)$. Similarly, $\pi_B(R) \subseteq Q$ for some $Q \in \syl_p(B)$. Thus, $R \subseteq P \times Q$. Since $|P \times Q| = |P||Q| = p^{\alpha + \beta}$, we must have $R = P \times Q$.

    To show $n_p(A \times B) = n_p(A)n_p(B)$, we note that the Sylow $p$-subgroups of $A \times B$ are precisely the subgroups of the form $P \times Q$, where $P \in \syl_p(A)$ and $Q \in \syl_p(B)$. Since there are $n_p(A)$ choices for $P$ and $n_p(B)$ choices for $Q$, we have $n_p(A \times B) = n_p(A)n_p(B)$.
\end{sol}

\begin{exercise}\label{ex5.1.5}
    Exhibit a nonnormal subgroup of $Q_8 \times Z_4$ (note that every subgroup of each factor is normal).
\end{exercise}

\begin{sol}
    Note that $Q_8$ is non-abelian, so we may use the idea of a diagonal subgroup from \hyperref[ex3.1.38]{Exercise 3.1.38}. In particular, let $H = \gen{(i, 1)} = \set{(1, 0), (i, 1), (-1, 2), (-i, 3)}$. Then $H$ is a subgroup of $Q_8 \times Z_4$ that is isomorphic to $\gen{i}$. However,
    \[(j, 0)(i, 1)(j, 0)\inv = (ji(-j), 0 + 1 + 0) = (-i, 1)\]
    is not in $H$, so $H$ is not normal in $Q_8 \times Z_4$.
\end{sol}

\begin{exercise}\label{ex5.1.6}
    Show that all subgroups of $Q_8 \times E_{2^n}$ are normal.
\end{exercise}

\begin{sol}
    Let $H \leq Q_8 \times E_{2^n}$. Suppose $(a, b) \in H$ and $(x, y) \in Q_8 \times E_{2^n}$. Note that $(a\inv, b) \in H$, since every $b \in E_{2^n}$ is its own inverse. Then
    \[(x, y)(a, b)(x, y)\inv = (xax\inv, yby\inv) = (xax\inv, b)\]
    since $E_{2^n}$ is abelian. If $a \in Z(Q_8)$, then $xax\inv = a$, so $(x, y)(a, b)(x, y)\inv = (a, b) \in H$. If $a \notin Z(Q_8)$, then $a$ is one of $i, j, k$ or their inverses. In all cases, we have $xax\inv$ is one of $i, j, k$ or their inverses, so $xax\inv \in H$. Thus, $(x, y)(a, b)(x, y)\inv \in H$ for all $(a, b) \in H$ and $(x, y) \in Q_8 \times E_{2^n}$, so $H$ is normal in $Q_8 \times E_{2^n}$.
\end{sol}

\begin{exercise}\label{ex5.1.7}
    Let $G_1, G_2, \ldots, G_n$ be groups and $\pi$ be a fixed element of $S_n$. Prove that the map 
    \[\phi_\pi : G_1 \times G_2 \times \cdots \times G_n \to G_{\pi\inv(1)} \times G_{\pi\inv(2)} \times \cdots \times G_{\pi\inv(n)}\]
    defined by
    \[\phi_\pi(g_1, g_2, \ldots, g_n) = (g_{\pi\inv(1)}, g_{\pi\inv(2)}, \ldots, g_{\pi\inv(n)})\]
    is an isomorphism (so that changing the order of the factors in a direct product does not change the isomorphism type).
\end{exercise}

\begin{sol}
    Let $\pi \in S_n$ and let $\phi_\pi$ be the map defined in the problem statement. Let $g = (g_1, g_2, \ldots, g_n)$ and $h = (h_1, h_2, \ldots, h_n)$ be elements of $G_1 \times G_2 \times \cdots \times G_n$. Then
    \begin{align*}
        \phi_\pi(gh) & = \phi_\pi(g_1h_1, g_2h_2, \ldots, g_nh_n) \\
        & = (g_{\pi\inv(1)}h_{\pi\inv(1)}, g_{\pi\inv(2)}h_{\pi\inv(2)}, \ldots, g_{\pi\inv(n)}h_{\pi\inv(n)}) \\
        & = (g_{\pi\inv(1)}, g_{\pi\inv(2)}, \ldots, g_{\pi\inv(n)})(h_{\pi\inv(1)}, h_{\pi\inv(2)}, \ldots, h_{\pi\inv(n)}) \\
        & = \phi_\pi(g)\phi_\pi(h)
    \end{align*}
    Noting that $\pi \in S_n$ is a bijection, consider the map $\phi_{\pi\inv} : G_{\pi\inv(1)} \times G_{\pi\inv(2)} \times \cdots \times G_{\pi\inv(n)} \to G_1 \times G_2 \times \cdots \times G_n$ defined by
    \[\phi_{\pi\inv}(g_1, g_2, \ldots, g_n) = (g_{\pi(1)}, g_{\pi(2)}, \ldots, g_{\pi(n)})\]
    Then $\phi_{\pi\inv}$ is a homomorphism by the same argument as above. Moreover, $\phi_{\pi\inv}$ is the inverse of $\phi_\pi$, since
    \[\phi_\pi(\phi_{\pi\inv}(g_1, g_2, \ldots, g_n)) = \phi_\pi(g_{\pi(1)}, g_{\pi(2)}, \ldots, g_{\pi(n)}) = (g_1, g_2, \ldots, g_n)\]
    and
    \[\phi_{\pi\inv}(\phi_\pi(g_1, g_2, \ldots, g_n)) = \phi_{\pi\inv}(g_{\pi\inv(1)}, g_{\pi\inv(2)}, \ldots, g_{\pi\inv(n)}) = (g_1, g_2, \ldots, g_n).\]
    Therefore, $\phi_\pi$ is an isomorphism.
\end{sol}

\begin{exercise}\label{ex5.1.8}
    Let $G_1 = G_2 = \cdots = G_n$ and let $G = G_1 \times \cdots \times G_n$. Under the notation of the preceding exercise show that $\phi_\pi \in \text{Aut}(G)$. Show also that the map $\pi \mapsto \phi_\pi$ is an injective homomorphism of $S_n$ into $\text{Aut}(G)$. (In particular, $\phi_{\pi_1} \circ \phi_{\pi_2} = \phi_{\pi_1 \pi_2}$. It is at this point that the $\pi^{-1}$'s in the definition of $\phi_\pi$ are needed. The underlying reason for this is because if $e_i$ is the $n$-tuple with 1 in position $i$ and zeros elsewhere, $1 \leq i \leq n$, then $S_n$ acts on $\{e_1, \ldots, e_n\}$ by $\pi \cdot e_i = e_{\pi(i)}$; this is a left group action. If the $n$-tuple $(g_1, \ldots, g_n)$ is represented by $g_1 e_1 + \cdots + g_n e_n$, then this left group action on $\{e_1, \ldots, e_n\}$ extends to a left group action on sums by
    \[\pi \cdot (g_1 e_1 + g_2 e_2 + \cdots + g_n e_n) = g_1 e_{\pi(1)} + g_2 e_{\pi(2)} + \cdots + g_n e_{\pi(n)}.\]
    The coefficient of $e_{\pi(i)}$ on the right hand side is $g_i$, so the coefficient of $e_i$ is $g_{\pi^{-1}(i)}$. Thus the right hand side may be rewritten as $g_{\pi^{-1}(1)} e_1 + g_{\pi^{-1}(2)} e_2 + \cdots + g_{\pi^{-1}(n)} e_n$, which is precisely the sum attached to the $n$-tuple $(g_{\pi^{-1}(1)}, g_{\pi^{-1}(2)}, \ldots, g_{\pi^{-1}(n)})$. In other words, any permutation of the ``position vectors'' $e_1, \ldots, e_n$ (which fixes their coefficients) is the same as the inverse permutation on the coefficients (fixing the $e_i$'s). If one uses $\pi$'s in place of $\pi^{-1}$'s in the definition of $\phi_\pi$ then the map $\pi \mapsto \phi_\pi$ is not necessarily a homomorphism --- it corresponds to a \textit{right} group action.)
\end{exercise}

\begin{sol}
    Note that $\phi_\pi \in \aut(G)$ since $\phi_\pi$ is an isomorphism from $G$ to itself. Let $\pi_1, \pi_2 \in S_n$. Then
    \begin{align*}
        \phi_{\pi_1}(\phi_{\pi_2}(g_1, g_2, \ldots, g_n)) & = \phi_{\pi_1}(g_{\pi_2\inv(1)}, g_{\pi_2\inv(2)}, \ldots, g_{\pi_2\inv(n)}) \\
        & = (g_{\pi_2\inv(\pi_1\inv(1))}, g_{\pi_2\inv(\pi_1\inv(2))}, \ldots, g_{\pi_2\inv(\pi_1\inv(n))}) \\
        & = (g_{(\pi_1 \pi_2)\inv(1)}, g_{(\pi_1 \pi_2)\inv(2)}, \ldots, g_{(\pi_1 \pi_2)\inv(n)}) \\
        & = \phi_{\pi_1 \pi_2}(g_1, g_2, \ldots, g_n)
    \end{align*}
    Therefore, the map $\pi \mapsto \phi_\pi$ is a homomorphism. To see that it is injective, note that if $\phi_\pi$ is the identity automorphism, then for all $i$, $g_{\pi\inv(i)} = g_i$ for all $g_i \in G_i$, which implies $\pi$ is the identity permutation. Hence, the map is injective.
\end{sol}

\begin{exercise}\label{ex5.1.9}
    Let $G_i$ be a field $F$ for all $i$ and use the preceding exercise to show that the set of $n \times n$ matrices with one 1 in each row and each column is a subgroup of $\gl_n(F)$ isomorphic to $S_n$ (these matrices are the \textit{permutation matrices} since they simply permute the standard basis $e_1, \ldots, e_n$ (as above) of the $n$-dimensional vector space $F \times F \times \cdots \times F$).
\end{exercise}

\begin{sol}
    Let $G = F \times F \times \cdots \times F$ ($n$ times) and let $\pi \in S_n$. Then $\phi_\pi$ is an automorphism of $G$ that permutes the standard basis vectors $e_1, e_2, \ldots, e_n$. In particular, $\phi_\pi(e_i) = e_{\pi(i)}$ for all $i$. Thus, the matrix representation of $\phi_\pi$ with respect to the standard basis is a permutation matrix. Since the map $\pi \mapsto \phi_\pi$ is an injective homomorphism from $S_n$ to $\aut(G)$, the set of permutation matrices forms a subgroup of $\gl_n(F)$ that is isomorphic to $S_n$.
\end{sol}

\begin{exercise}\label{ex5.1.10}
    Let $p$ be a prime. Let $A$ and $B$ be two cyclic groups of order $p$ with generators $x$ and $y$ respectively. Set $E = A \times B$ so that $E$ is the elementary abelian group of order $p^2$: $E_{p^2}$ Prove that the distinct subgroups of $E$ of order $p$ are
    \[\langle x \rangle, \quad \langle xy \rangle, \quad \langle xy^2 \rangle, \quad \ldots, \quad \langle xy^{p-1} \rangle, \quad \langle y \rangle\] (note that there are $p + 1$ of them).
\end{exercise}

\begin{sol}
    Note that any proper, nontrivial subgroup of $E$ must have order $p$, and every element of $E$ has order $p$. The text shows that there are $p + 1$ subgroups of $E$ of order $p$, so it suffices to show that the subgroups listed in the problem statement are distinct. Note that $\gen x$ and $\gen y$ are distinct since they have different generators. Moreover, $\gen{xy^i}$ is distinct from $\gen x$ and $\gen y$ for all $i$, since $xy^i$ is not a power of $x$ or $y$ (as can be seen because $\gen x \cap \gen y$ is trivial). Finally, $\gen{xy^i}$ is distinct from $\gen{xy^j}$ for $i \neq j$, since if $\gen{xy^i} = \gen{xy^j}$, then $xy^i$ is a power of $xy^j$, which implies $y^{i-j}$ is a power of $x$, which is impossible since $\gen x \cap \gen y$ is trivial. Thus, all the subgroups listed in the problem statement are distinct.
\end{sol}

\begin{exercise}\label{ex5.1.11}
    Let $p$ be a prime, and let $n \in \zp$. Find a formula for the number of subgroups of order $p$ in the elementary abelian group $E_{p^n}$.
\end{exercise}

\begin{sol}
    Every element of $E_{p^n}$ has order $p$, so every subgroup of order $p$ is generated by a single non-identity element. Since there are $p^n - 1$ non-identity elements in $E_{p^n}$, and each subgroup of order $p$ has $p - 1$ generators (the non-identity elements of the subgroup), the number of subgroups of order $p$ in $E_{p^n}$ is
    \[\frac{p^n - 1}{p - 1} = 1 + p + p^2 + \cdots + p^{n-1}\]
    as desired.
\end{sol}

\begin{specialexercise}\label{ex5.1.12}
    Let $A$ and $B$ be groups. Assume $Z(A)$ contains a subgroup $Z_1$ and $Z(B)$ contains a subgroup $Z_2$ with $Z_1 \cong Z_2$. Let this isomorphism be given by the map $x_i \mapsto y_i$ for all $x_i \in Z_1$. A \textbf{\textit{central product}} of $A$ and $B$ is a quotient
    \[(A \times B)/Z \quad \text{where} \quad Z = \{(x_i, y_i^{-1}) \mid x_i \in Z_1\}\]
    and is denoted by $A * B$ --- it is not unique since it depends on $Z_1$, $Z_2$ and the isomorphism between them. (Think of $A * B$ as the direct product of $A$ and $B$ ``collapsed'' by identifying each element $x_i \in Z_1$ with its corresponding element $y_i \in Z_2$.)
    \begin{subproblems}
        \item Prove that the images of $A$ and $B$ in the quotient group $A * B$ are isomorphic to $A$ and $B$, respectively, and that these images intersect in a central subgroup isomorphic to $Z_1$. Find $|A * B|$.
        \item Let $Z_4 = \langle x \rangle$. Let $D_8 = \langle r, s \rangle$ and $Q_8 = \langle i, j \rangle$ be given by their usual generators and relations. Let $Z_4 * D_8$ be the central product of $Z_4$ and $D_8$ which identifies $x^2$ and $r^2$ (i.e., $Z_1 = \langle x^2 \rangle$, $Z_2 = \langle r^2 \rangle$ and the isomorphism is $x^2 \mapsto r^2$) and let $Z_4 * Q_8$ be the central product of $Z_4$ and $Q_8$ which identifies $x^2$ and $-1$. Prove that $Z_4 * D_8 \cong Z_4 * Q_8$.
    \end{subproblems}
\end{specialexercise}

\begin{solalph}
    \item Consider the maps $\pi_A : A \to A * B$ and $\pi_B : B \to A * B$ defined by $\pi_A(a) = (a, 1)Z$ and $\pi_B(b) = (1, b)Z$. It is clear that $\pi_A$ and $\pi_B$ are homomorphisms. Moreover, $\ker\pi_A$ consists of the elements $a \in A$ such that $(a, 1) \in Z$. Then $y_i = 1$, and since $x_i \mapsto y_i$ is an isomorphism, we have $x_i = 1$, so $a = 1$. Thus, $\ker\pi_A$ is trivial, and $\pi_A$ is injective. Similarly, $\ker\pi_B$ is trivial, and $\pi_B$ is injective. Therefore, the images of $A$ and $B$ in $A * B$ are isomorphic to $A$ and $B$, respectively. Moreover, the images of $A$ and $B$ in $A * B$ intersect in the subgroup $\{(x_i, 1)Z \mid x_i \in Z_1\}$, which is isomorphic to $Z_1$. Finally, since $Z$ has $|Z_1|$ elements, we have
    \[|A * B| = \frac{|A||B|}{|Z_1|}.\]
    \item Let $Z_D$ denote the $Z$ in the quotient for $Z_4 * D_8$ and let $Z_Q$ denote the $Z$ in the quotient for $Z_4 * Q_8$. Let $x$ generate $Z_4$ in $Z_4 * D_8$ and $y$ generate $Z_4$ in $Z_4 * Q_8$. Let $r$ and $s$ be the generators of $D_8$ in $Z_4 * D_8$ and let $i$ and $j$ be the generators of $Q_8$ in $Z_4 * Q_8$. Let $\b x$ denote the coset $(x, 1)Z_D$ in $Z_4 * D_8$, and use similar notation for the elements $r, s, y, i, j$. Let $\b z = \b j\b y\inv$. Consider the map $\phi : Z_4 * D_8 \to Z_4 * Q_8$ defined by 
    \[\phi(\b x^\alpha\b r^\beta\b s^\gamma) = \b y^\alpha\b i^\beta\b z^\gamma.\]
    To see that this is well-defined, suppose $\b x^\alpha\b r^\beta\b s^\gamma = \b 1$ in $Z_4 * D_8$. Then $\b r^\beta\b s^\gamma$ is in the image of $Z_4$ in $Z_4 * D_8$, so $\b r^\beta\b s^\gamma = \b x^{2\delta}$ for some $\delta$. Since $r^2$ is central in $D_8$, we have $\b r^\beta\b s^\gamma = \b r^{\beta + 2\delta}\b s^\gamma$. If $\gamma = 0$, then $\b r^{\beta + 2\delta} = \b 1$, so $4 \mid \beta + 2\delta$. If $\gamma = 1$, then $\b r^{\beta + 2\delta}\b s = \b 1$, so $4 \mid \beta + 2\delta - 2$. In either case, we have $4 \mid \beta + 2\delta$, so $\b r^\beta\b s^\gamma = \b x^{2\delta} = \b 1$. Thus, $\alpha$ is even, and we have $\phi(\b x^\alpha\b r^\beta\b s^\gamma) = \b y^\alpha\b i^\beta\b z^\gamma = \b 1$. Therefore, $\phi$ is well-defined. It is clear that $\phi$ is a homomorphism. Moreover, if $\phi(\b x^\alpha\b r^\beta\b s^\gamma) = \b 1$, then $\alpha$ is even and $\beta$ and $\gamma$ are zero, so $\phi$ is injective. Finally, since $|Z_4 * D_8| = |Z_4 * Q_8|$, we conclude that $\phi$ is an isomorphism.
\end{solalph}

\begin{exercise}\label{ex5.1.13}
    Give presentations for the groups $Z_4 * D_8$ and $Z_4 * Q_8$ constructed in the preceding exercise.
\end{exercise}

\begin{sol}
    To construct the presentations, we note that $\b x$ should satisfy the relations of $Z_4$, $\b r$ and $\b s$ should satisfy the relations of $D_8$, and $\b i$ and $\b z$ should satisfy the relations of $Q_8$. Moreover, we have the relations $\b x^2 = \b r^2$ in $Z_4 * D_8$, since $\b x^2\b r^{-2} = \b 1$. Lastly, $\b x$ is central in that $\b x \b r = \b r\b x$ and $\b x\b s = \b s\b x$ in $Z_4 * D_8$. Then a presentation for $Z_4 * D_8$ is
    \[\gen{\b x, \b r, \b s \mid \b r^4 = \b s^2 = \b 1, \b x^2 = \b r^2, \b s\b r\b s = \b r\inv, \b x\b r = \b r\b x, \b x\b s = \b s\b x}.\]
    To find a presentation for $Z_4 * Q_8$, we may use the isomorphism $\phi$ from the preceding exercise to rewrite the presentation for $Z_4 * D_8$ in terms of $\b y, \b i, \b z$. In particular, we have $\b y^2 = \b i^2\b z^2$, $\b z^2 = \b 1$, and $\b y\b i = \b i\b y\b z$. Thus, a presentation for $Z_4 * Q_8$ is
    \[\gen{\b y, \b i, \b z \mid \b i^4 = \b z^2 = \b 1, \b y^2 = \b i^2, \b z\b i \b z = \b i\inv, \b y\b i = \b i\b y, \b y\b z = \b z\b y}. \qh\]
\end{sol}

\begin{exercise}\label{ex5.1.14}
    Let $G = A_1 \times A_2 \times \cdots \times A_n$, and for each $i$, let $B_i$ be a normal subgroup of $A_i$. Prove that $B_1 \times B_2 \times \cdots \times B_n \nsub G$, and that
    \[(A_1 \times A_2 \times \cdots \times A_n)/(B_1 \times B_2 \times \cdots \times B_n) \cong (A_1/B_1) \times (A_2/B_2) \times \cdots \times (A_n/B_n).\]
\end{exercise}

\begin{sol}
    Let $G = A_1 \times A_2 \times \cdots \times A_n$ and let $B_i \nsub A_i$ for all $i$. Let $H = B_1 \times B_2 \times \cdots \times B_n$. It is clear that $H$ is a subgroup of $G$. Moreover, since each $B_i$ is normal in $A_i$, we have that for any $g = (a_1, a_2, \ldots, a_n) \in G$ and any $h = (b_1, b_2, \ldots, b_n) \in H$, we have
    \[ghg\inv = (a_1b_1a_1\inv, a_2b_2a_2\inv, \ldots, a_nb_na_n\inv) \in H\]
    since $a_ib_ia_i\inv \in B_i$ for all $i$. Thus, $H$ is normal in $G$.

    Consider the map $\phi : G \to (A_1/B_1) \times (A_2/B_2) \times \cdots \times (A_n/B_n)$ defined by
    \[\phi((a_1, a_2, \ldots, a_n)) = (a_1B_1, a_2B_2, \ldots, a_nB_n).\]
    It is clear that $\phi$ is a homomorphism. Moreover, $\phi$ is surjective since for any $(a_1B_1, a_2B_2, \ldots, a_nB_n)$, we have $\phi((a_1, a_2, \ldots, a_n)) = (a_1B_1, a_2B_2, \ldots, a_nB_n)$. Finally, the kernel of $\phi$ is the set of all $(a_1, a_2, \ldots, a_n)$ such that $a_i \in B_i$ for all $i$, which is exactly $H$. Thus, by the First Isomorphism Theorem, we have
    \[G/H \cong (A_1/B_1) \times (A_2/B_2) \times \cdots \times (A_n/B_n)\]
    as desired.
\end{sol}

\begin{exercise}\label{ex5.1.15}
    Let $I$ be any nonempty index set and let $(G_i, \star_i)$ be a group for each $i \in I$. The \textit{direct product} of the groups $G_i$, $i \in I$ is the set $G = \prod_{i \in I} G_i$ (the Cartesian product of the $G_i$'s) with a binary operation defined as follows: if $\prod a_i$ and $\prod b_i$ are elements of $G$, then their product in $G$ is given by
    \[\left(\prod_{i \in I} a_i\right)\left(\prod_{i \in I} b_i\right) = \prod_{i \in I} (a_i \star_i b_i)\]
    (i.e., the group operation in the direct product is defined componentwise).
    \begin{subproblems}
        \item Show that this binary operation is well defined and associative.
        \item Show that the element $\prod 1_i$ satisfies the axiom for the identity of $G$, where $1_i$ is the identity of $G_i$ for all $i$.
        \item Show that the element $\prod a_i^{-1}$ is the inverse of $\prod a_i$, where the inverse of each component element $a_i$ is taken in the group $G_i$.
    \end{subproblems}
    Conclude that the direct product is a group. (Note that if $I = \set{1, 2, \ldots, n}$, this definition of the direct product is the same as the $n$-tuple definition in the text.)
\end{exercise}

\begin{solalph}
    \item To show well-definedness, let $\prod \alpha_i = \prod a_i$ and $\prod \beta_i = \prod b_i$. Then $\alpha_i = a_i$ and $\beta_i = b_i$ for all $i$, so
    \[\left(\prod_{i \in I} \alpha_i\right)\left(\prod_{i \in I} \beta_i\right) = \prod_{i \in I} (\alpha_i \star_i \beta_i) = \prod_{i \in I} (a_i \star_i b_i)\]
    as desired. To show associativity, let $\prod a_i$, $\prod b_i$, and $\prod c_i$ be elements of $G$. Then
    \begin{align*}
        \left(\prod_{i \in I} a_i\right)\left(\left(\prod_{i \in I} b_i\right)\left(\prod_{i \in I} c_i\right)\right) & = \left(\prod_{i \in I} a_i\right)\left(\prod_{i \in I} (b_i \star_i c_i)\right) \\
        & = \prod_{i \in I} (a_i \star_i (b_i \star_i c_i)) \\
        & = \prod_{i \in I} ((a_i \star_i b_i) \star_i c_i) \\
        & = \left(\prod_{i \in I} (a_i \star_i b_i)\right)\left(\prod_{i \in I} c_i\right) \\
        & = \left(\left(\prod_{i \in I} a_i\right)\left(\prod_{i \in I} b_i\right)\right)\left(\prod_{i \in I} c_i\right)
    \end{align*}
    as desired.
    \item Let $\prod a_i$ be an element of $G$. Then
    \[\left(\prod_{i \in I} 1_i\right)\left(\prod_{i \in I} a_i\right) = \prod_{i \in I} (1_i \star_i a_i) = \prod_{i \in I} a_i\]
    and
    \[\left(\prod_{i \in I} a_i\right)\left(\prod_{i \in I} 1_i\right) = \prod_{i \in I} (a_i \star_i 1_i) = \prod_{i \in I} a_i.\]
    Thus, $\prod 1_i$ is the identity element of $G$.
    \item Let $\prod a_i$ be an element of $G$. Then
    \[\left(\prod_{i \in I} a_i\right)\left(\prod_{i \in I} a_i^{-1}\right) = \prod_{i \in I} (a_i \star_i a_i^{-1}) = \prod_{i \in I} 1_i\]
    and
    \[\left(\prod_{i \in I} a_i^{-1}\right)\left(\prod_{i \in I} a_i\right) = \prod_{i \in I} (a_i^{-1} \star_i a_i) = \prod_{i \in I} 1_i.\]
    Thus, $\prod a_i^{-1}$ is the inverse of $\prod a_i$. We conclude that $G$ is a group with inverse and identity as described above.
\end{solalph}

\begin{exercise}\label{ex5.1.16}
    State and prove the generalization of Proposition 2 to arbitrary direct products.
\end{exercise}

\begin{sol}
    Proposition 2 stated for direct arbitrary groups is as follows: if $G = \prod_{i \in I} G_i$ is a direct product of groups, then
    \begin{enumerate}
        \item Fix each $i$ with $j$-th coordinate where $j \neq i$ is the identity element of $G_j$ for $1 \leq j \leq n$. Then
        \[\wh G_i = \set{(1, 1, \ldots, 1, g_i, 1, \ldots) \mid g_i \in G_i}\] is a normal subgroup of $G$ that is isomorphic to $G_i$, or that
        \[G/\wh G_i \cong \prod_{j \neq i} G_j.\]
        \item For fixed $i$, define $\pi_i : G \to G_i$ by $\pi_i((g_j)_{j \in I}) = g_i$. Then $\pi_i$ is a surjective homomorphism with kernel $\wh G_i$, so $G/\wh G_i \cong G_i$.
        \item If $x \in G_i$ and $y \in G_j$ for $i \neq j$, then $x$ and $y$ commute in $G$.
    \end{enumerate}
    The proof is similar to the proof of Proposition 2 in the text.
    \begin{enumerate}
        \item It is clear that $\wh G_i$ is a subgroup of $G$. Moreover, for any $\prod_{j \in I} g_j \in G$ and any $h = \prod_{j \neq i} 1_j h_i \in \wh G_i$, we have
        \[\left(\prod_{j \in I} g_j\right)h\left(\prod_{j \in I} g_j\right)\inv = \prod_{j \neq i} 1_j (g_i h_i g_i\inv) \in \wh G_i\]
        since $g_i h_i g_i\inv \in G_i$. Thus, $\wh G_i$ is normal in $G$. Consider the map $\phi : G_i \to \wh G_i$ defined by $\phi(g_i) = \prod_{j \neq i} 1_j g_i$. It is clear that $\phi$ is a homomorphism. Moreover, if $\phi(g_i) = \prod_{j \neq i} 1_j g_i = \prod_{j \neq i} 1_j$, then $g_i$ is the identity element of $G_i$, so $\phi$ is injective. Finally, for any $\prod_{j \neq i} 1_j h_i \in \wh G_i$, we have $\phi(h_i) = \prod_{j \neq i} 1_j h_i$, so $\phi$ is surjective. Thus, $\wh G_i$ is isomorphic to $G_i$. 
        
        Consider the map
        \[\psi : G \to \prod_{j \neq i} G_j \text{ defined by } \psi\left(\prod_{j \in I} g_j\right) = \prod_{j \neq i} g_j.\]
        It is clear that $\psi$ is a homomorphism. Moreover, the kernel of $\psi$ is $\wh G_i$, so by the First Isomorphism Theorem, we have
        \[G/\wh G_i \cong \prod_{j \neq i} G_j.\]
        \item For fixed $i$, it is clear that $\pi_i$ is a homomorphism. Moreover, $\pi_i$ is surjective since for any $g_i \in G_i$, we have $\pi_i(\prod_{j \in I} 1_j g_i) = g_i$. Finally, the kernel of $\pi_i$ is $\wh G_i$, so by the First Isomorphism Theorem, we have
        \[G/\wh G_i \cong G_i.\]
        \item Let $x = \prod_{j \neq i} 1_j x_i$ and $y = \prod_{j \neq k} 1_j y_k$ for some $i \neq k$. Then
        \[xy = \prod_{j \neq i, k} 1_j x_i y_k = \prod_{j \neq i, k} 1_j y_k x_i = yx\]
        as desired.
    \end{enumerate}
\end{sol}

\begin{exercise}\label{ex5.1.17}
    Let $I$ be any nonempty index set and let $G_i$ be a group for each $i \in I$. The \textit{restricted direct product} or \textit{direct sum} of the groups $G_i$ is the set of elements of the direct product which are the identity in all but finitely many components, that is, the set of all elements $\prod a_i \in \prod_{i \in I} G_i$ such that $a_i = 1_i$ for all but a finite number of $i \in I$.
    \begin{subproblems}
        \item Prove that the restricted direct product is a subgroup of the direct product.
        \item Prove that the restricted direct product is normal in the direct product.
        \item Let $I = \mathbb{Z}^+$ and let $p_i$ be the $i$th integer prime. Show that if $G_i = \mathbb{Z}/p_i\mathbb{Z}$ for all $i \in \mathbb{Z}^+$, then every element of the restricted direct product of the $G_i$'s has finite order but $\prod_{i \in \mathbb{Z}^+} G_i$ has elements of infinite order. Show that in this example the restricted direct product is the torsion subgroup of the direct product (cf. \hyperref[ex2.1.16]{Exercise 2.1.16}).
    \end{subproblems}
\end{exercise}

\begin{solalph}
    \item Let $G = \prod_{i \in I} G_i$ and let $H$ be the restricted direct product of the $G_i$'s. It is clear that $H$ is a subset of $G$. Moreover, if $\prod a_i$ and $\prod b_i$ are elements of $H$, then there are only finitely many $i$ such that $a_i \neq 1_i$ and only finitely many $i$ such that $b_i \neq 1_i$. Thus, there are only finitely many $i$ such that $a_i b_i \neq 1_i$, so $\prod a_i \prod b_i = \prod (a_i b_i)$ is an element of $H$. Finally, if $\prod a_i$ is an element of $H$, then there are only finitely many $i$ such that $a_i \neq 1_i$, so there are only finitely many $i$ such that $a_i^{-1} \neq 1_i$, so $\prod a_i^{-1}$ is an element of $H$. Thus, $H \leq G$.
    \item Let $G = \prod_{i \in I} G_i$ and let $H$ be the restricted direct product of the $G_i$'s. Let $\prod a_i$ be an element of $G$ and let $\prod b_i$ be an element of $H$. Then there are only finitely many $i$ such that $b_i \neq 1_i$, so there are only finitely many $i$ such that $a_i b_i a_i^{-1} \neq 1_i$, so $\prod a_i \prod b_i \prod a_i^{-1}$ is an element of $H$. Thus, $H \nsub G$.
    \item Let $G = \prod_{i \in \mathbb{Z}^+} G_i$ where $G_i = \mathbb{Z}/p_i\mathbb{Z}$ for all $i$. Let $H$ be the restricted direct product of the $G_i$'s. If $\prod a_i$ is an element of $H$, then there are only finitely many $i$ such that $a_i \neq 1_i$, so there are only finitely many $i$ such that $a_i$ has order greater than 1, so $\prod a_i$ has finite order. On the other hand, if $\prod a_i$ is an element of $G$ such that $a_i$ is a generator of $G_i$ for all $i$, then $\prod a_i$ has infinite order since for any positive integer $n$, we have $(\prod a_i)^n = \prod a_i^n$, and there are infinitely many $i$ such that $a_i^n \neq 1_i$.
    
    To show $H = \tor(G)$, we note that $H \subseteq \tor(G)$ since every element of $H$ has finite order. Conversely, if $\prod a_i$ is an element of $\tor(G)$, then there is some positive integer $n$ such that $(\prod a_i)^n = \prod a_i^n = \prod 1_i$, so $a_i^n = 1_i$ for all $i$. Since $G_i$ is cyclic of prime order, this implies that $a_i$ is either the identity or a generator of $G_i$. If there are infinitely many $i$ such that $a_i$ is a generator of $G_i$, then $\prod a_i$ has infinite order, which is a contradiction. Thus, there are only finitely many $i$ such that $a_i$ is a generator of $G_i$, so there are only finitely many $i$ such that $a_i \neq 1_i$, so $\prod a_i$ is an element of $H$. Therefore, $\tor(G) \subseteq H$, and we conclude that $\tor(G) = H$.
\end{solalph}

\begin{exercise}\label{ex5.1.18}
    In each of (a) to (e) give an example of a group with the specified properties:
    \begin{subproblems}
        \item an infinite group in which every element has order 1 or 2
        \item an infinite group in which every element has finite order but for each positive integer $n$ there is an element of order $n$
        \item a group with an element of infinite order and an element of order 2
        \item a group $G$ such that every finite group is isomorphic to some subgroup of $G$
        \item a nontrivial group $G$ such that $G \cong G \times G$.
    \end{subproblems}
\end{exercise}

\begin{solalph}
    \item Consider the infinite direct product $\intmod[2] \times \intmod[2] \times \cdots$. Every element of this group is a tuple of 0's and 1's, and the order of each element is either 1 (if the tuple is the identity) or 2 (if the tuple has at least one 1).
    \item Let $G = \intmod[2] \times \intmod[3] \times \intmod[4] \times \cdots$. Every element of $G$ has finite order since it is a tuple of elements from finite cyclic groups. Moreover, for each positive integer $n$, the element that is the identity in all components except the $n$-th component, which is a generator of $\intmod[n]$, has order $n$.
    \item Consider the group $\intmod[2] \times \mathbb{Z}$. The element $(1, 0)$ has order 2, and the element $(0, 1)$ has infinite order.
    \item By Cayley's Theorem, every finite group is isomorphic to a subgroup of some symmetric group $S_n$. We may then consider the group $G = S_1 \times S_2 \times S_3 \times \cdots$. For any finite group $H$, there is some $n$ such that $H$ is isomorphic to a subgroup of $S_n$, so $H$ is isomorphic to a subgroup of $G$.
    \item Let $G = \z \times \z \times \cdots$. Consider the map $\phi : G \to G \times G$ defined by
    \[\phi((a_1, a_2, a_3, \ldots)) = ((a_1, a_3, a_5, \ldots), (a_2, a_4, a_6, \ldots)).\]
    It is clear that $\phi$ is a homomorphism. Moreover, if $\phi((a_1, a_2, a_3, \ldots)) = ((0, 0, 0, \ldots), (0, 0, 0, \ldots))$, then $a_i = 0$ for all $i$, so $\phi$ is injective. Finally, let $((b_1, b_2, b_3, \ldots), (c_1, c_2, c_3, \ldots))$ be an element of $G \times G$. Then $\phi((b_1, c_1, b_2, c_2, b_3, c_3, \ldots)) = ((b_1, b_2, b_3, \ldots), (c_1, c_2, c_3, \ldots))$, so $\phi$ is surjective. Thus, $\phi$ is an isomorphism and $G \cong G \times G$.
\end{solalph}

\newpage

\subsection{The Fundamental Theorem of Finitely Generated Abelian Groups}

\begin{exercise}\label{ex5.2.1}
    In each parts (a) to (e), give the number of nonisomorphic abelian groups of the specified order---do not list the groups:

    \noindent (a) order 100, \quad (b) order 576, \quad (c) order 1155, \quad (d) order 42875, \quad (e) order 2704.
\end{exercise}

\begin{solalph}
    \item $100 = 2^2 \cdot 5^2$; 2 partitions of 2, hence $2 \cdot 2 = 4$ nonisomorphic abelian groups of order 100.
    \item $576 = 2^6 \cdot 3^2$; 11 partitions of 6, hence $11 \cdot 2 = 22$ nonisomorphic abelian groups of order 576.
    \item $1155 = 3 \cdot 5 \cdot 7 \cdot 11$. Only one abelian group of order 1155.
    \item $42875 = 5^3 \cdot 7^3$; 3 partitions of 3, hence $3 \cdot 3 = 9$ nonisomorphic abelian groups of order 42875.
    \item $2704 = 2^4 \cdot 13^2$; 5 partitions of 4, hence $5 \cdot 2 = 10$ nonisomorphic abelian groups of order 2704.
\end{solalph}

\begin{exercise}\label{ex5.2.2}
    In each of parts (a) to (e), give the lists of invariant factors for all abelian groups of the specified order:

    \noindent (a) group 270, \quad (b) order 9801, \quad (c) order 320, \quad (d) order 105, \quad (e) order 44100.
\end{exercise}

\begin{solalph}
    \item $270 = 2 \cdot 3^3 \cdot 5$. The invariant factors are $\set{270}, \set{90, 3}, \set{30, 3, 3}$.
    \item $9801 = 3^4 \cdot 11^2$. The invariant factors are $\set{9801}, \set{891, 11}, \set{3267, 3}, \set{297, 33}, \set{1089, 9}, \set{99, 99}, \\
    \set{1089, 3, 3}, \set{99, 33, 3}, \set{363, 3, 3, 3}, \set{33, 33, 3, 3}$.
    \item $320 = 2^6 \cdot 5$. The invariant factors are $\set{320}, \set{160, 2}, \set{80, 4}, \set{80, 2, 2}, \set{40, 8}, \set{40, 4, 2}, \set{40, 2, 2, 2}, \\
    \set{20, 4, 4}, \set{20, 4, 2, 2}, \set{20, 2, 2, 2, 2}, \set{10, 2, 2, 2, 2, 2}$.
    \item $105 = 3 \cdot 5 \cdot 7$. The invariant factors are $\set{105}$.
    \item $44100 = 2^2 \cdot 3^2 \cdot 5^2 \cdot 7^2$. The invariant factors are $\set{44100}, \set{22050, 2}, \set{14700, 3}, \set{8820, 5}, \set{6300, 7}, \set{4410, 10}, \\
    \set{3150, 14}, \set{2205, 15}, \set{1800, 21}, \set{1260, 35}, \set{1470, 30}, \set{1050, 42}, \set{630, 70}, \set{420, 105}, \set{210, 210}$.
\end{solalph}

\begin{exercise}\label{ex5.2.3}
    In each of the parts (a) to (e), give the lists of elementary divisors for all abelian groups of the specified order, and then match each list with the corresponding list of invariant factors found in the preceding exercise:

    \noindent (a) group 270, \quad (b) order 9801, \quad (c) order 320, \quad (d) order 105, \quad (e) order 44100.
\end{exercise}

\begin{solalph}
    \item $\set{2, 27, 5}, \set{2, 9, 5, 3}, \set{2, 3, 5, 3, 3}$.
    \item $\set{81, 121}, \set{81, 11, 11}, \set{27, 121, 3}, \set{27, 11, 3, 11}, \set{9, 121, 9}, \set{9, 11, 9, 11}, \set{9, 121, 3, 3}, \set{9, 11, 3, 11, 3}, \\
    \set{3, 121, 3, 3, 3}, \set{3, 11, 3, 11, 3, 3}$.
    \item $\set{64, 5}, \set{32, 5, 2}, \set{16, 5, 4}, \set{16, 5, 2, 2}, \set{8, 5, 8}, \set{8, 5, 4, 2}, \set{8, 5, 2, 2, 2}, \set{4, 5, 4, 4}, \set{4, 5, 4, 2, 2}, \\
    \set{4, 5, 2, 2, 2, 2}, \set{2, 5, 2, 2, 2, 2, 2}$.
    \item $\set{3, 5, 7}$.
    \item $\set{4, 9, 25, 49}, \set{2, 9, 25, 49, 2}, \set{4, 3, 25, 49, 3}, \set{4, 9, 5, 49, 5}, \set{4, 9, 25, 7, 7}, \set{2, 3, 25, 49, 2, 3}, \set{2, 9, 5, 49, 2, 5}, \\
    \set{2, 9, 25, 7, 2, 7}, \set{4, 3, 5, 49, 3, 5}, \set{2, 3, 25, 49, 2, 3}, \set{2, 9, 5, 49, 2, 7}, \set{4, 3, 25, 7, 3, 7}, \set{2, 3, 5, 49, 2, 3, 5}, \\
    \set{2, 3, 25, 7, 2, 3, 7}, \set{4, 3, 5, 7, 3, 5, 7}, \set{2, 3, 5, 7, 2, 3, 5, 7}$.
\end{solalph}

\begin{exercise}\label{ex5.2.4}
    In each of parts (a) to (d), determine which pairs of abelian groups listed are isomorphic (here the expression $\set{a_1, a_2, \ldots, a_k}$ denotes the abelian group $Z_{a_1} \times Z_{a_2} \times \cdots \times Z_{a_k}$).
    \begin{subproblems}
        \item $\{4, 9\}$, \quad $\{6, 6\}$, \quad $\{8, 3\}$, \quad $\{9, 4\}$, \quad $\{6, 4\}$, \quad $\{64\}$.
        \item $\{2^2, 2 \cdot 3^2\}$, \quad $\{2^2 \cdot 3, 2 \cdot 3\}$, \quad $\{2^3 \cdot 3^2\}$, \quad $\{2^2 \cdot 3^2, 2\}$.
        \item $\{5^2 \cdot 7^2,\ 3^2 \cdot 5 \cdot 7\}$, \quad $\{3^2 \cdot 5^2 \cdot 7,\ 5 \cdot 7^2\}$, \quad $\{3 \cdot 5^2,\ 7^2,\ 3 \cdot 5 \cdot 7\}$,
        \quad $\{5^2 \cdot 7,\ 3^2 \cdot 5,\ 7^2\}$.
        \item $\{2^2 \cdot 5 \cdot 7,\ 2^3 \cdot 5^3,\ 2 \cdot 5^2\}$, \quad $\{2^3 \cdot 5^3 \cdot 7,\ 2^3 \cdot 5^3\}$, \quad $\{2^2,\ 2 \cdot 7,\ 2^3,\ 5^3,\ 5^3\}, \quad \{2 \cdot 5^3,\ 2^2 \cdot 5^3,\ 2^3,\ 7\}$.
    \end{subproblems}
\end{exercise}

\begin{solalph}
    \item $\set{4, 9}$ and $\set{9, 4}$ have the same elementary divisors, so they are isomorphic. In order, the other four groups have elementary divisors $\set{3, 2, 3, 2}$, $\set{8, 3}$, $\set{3, 2, 4}$, and $\set{64}$, so they are not isomorphic to each other or to $\set{4, 9}$.
    \item $\set{2^2, 2 \cdot 3^2}$ and $\set{2^2 \cdot 3^2, 2}$ have the same elementary divisors, so they are isomorphic. In order, the other two groups have elementary divisors $\set{2^2 \cdot 3, 2 \cdot 3}$ and $\set{2^3 \cdot 3^2}$, so they are not isomorphic to each other or to $\set{2^2, 2 \cdot 3^2}$.
    \item The first, second, and fourth groups have the same elementary divisors $\set{5^2, 7^2, 3^2, 5, 7}$, so they are isomorphic. The third group has elementary divisors $\set{3, 5^2, 7^2, 3, 5, 7}$, so it is not isomorphic to the other three groups.
    \item The third and fourth groups have the elementary divisors $\set{2, 125, 4, 125, 8, 7}$, so they are isomorphic.
\end{solalph}

\begin{exercise}\label{ex5.2.5}
    Let $G$ be a finite abelian group of type $(n_1, n_2, \ldots, n_t)$. Prove that $G$ contains an element of order $m$ if and only if $m \mid n_1$. Deduce that $G$ is of exponent $n_1$.
\end{exercise}

\begin{sol}
    \leavevmode
    \begin{ifandonlyif}
        \item[\rightimp] Let $g = (g_1, g_2, \ldots, g_t) \in G$ have order $m$. Then $m = \lcm(|g_1|, |g_2|, \ldots, |g_t|)$, so $|g_i| \mid m$ for all $i$. Since $G$ is of type $(n_1, n_2, \ldots, n_t)$, we have $|g_i| \mid n_i$ for all $i$, so $m$ divides $\lcm(n_1, n_2, \ldots, n_t) = n_1$.
        \item[\leftimp] If $m \mid n_1$, then there exists an element of order $m$ in the cyclic subgroup of order $n_1$.
    \end{ifandonlyif}
    Since $G$ contains an element of order $n_1$, the exponent of $G$ is at least $n_1$. On the other hand, if $g \in G$, then $|g| \mid n_1$, so the exponent of $G$ divides $n_1$. Thus, the exponent of $G$ is exactly $n_1$.
\end{sol}

\begin{exercise}\label{ex5.2.6}
    Prove that any finite group has a finite exponent. Give an example of an infinite group with finite exponent. Does a finite group of exponent $m$ always contain an element of order $m$?
\end{exercise}

\begin{sol}
    Let $G$ be a finite group. Then there are only finitely many elements of $G$, so there are only finitely many orders of elements in $G$. Let $m$ be the least common multiple of the orders of all elements of $G$. Then for any $g \in G$, we have $g^m = e$, so $G$ has exponent dividing $m$.
    
    An example of an infinite group with finite exponent is the infinite direct product $\intmod[2] \times \intmod[2] \times \cdots$. Every element of this group has order 1 or 2, so the exponent of this group is 2.
    
    A finite group of exponent $m$ does not always contain an element of order $m$: consider $S_3$, which has exponent 6 but does not contain an element of order 6.
\end{sol}

\begin{exercise}\label{ex5.2.7}
    Let $p$ be a prime and let $A = \langle x_1 \rangle \times \langle x_2 \rangle \times \cdots \times \langle x_n \rangle$ be an abelian $p$-group, where $|x_i| = p^{\alpha_i} > 1$ for all $i$. Define the $p^{\text{th}}$-\textit{power map}
    \[\phi : A \to A \quad \text{by} \quad \phi : x \mapsto x^p.\]
    \begin{subproblems}
        \item Prove that $\phi$ is a homomorphism.
        \item Describe the image and kernel of $\phi$ in terms of the given generators.
        \item Prove both $\ker \phi$ and $A / \operatorname{im} \phi$ have rank $n$ (i.e., have the same rank as $A$) and prove these groups are both isomorphic to the elementary abelian group, $E_{p^n}$, of order $p^n$.
    \end{subproblems}
\end{exercise}

\begin{solalph}
    \item Let $x = (x_1^{s_1}, x_2^{s_2}, \ldots, x_n^{s_n})$ and $y = (x_1^{t_1}, x_2^{t_2}, \ldots, x_n^{t_n})$ be elements of $A$. Then
    \begin{align*}
        \phi(xy) & = \phi((x_1^{s_1 + t_1}, x_2^{s_2 + t_2}, \ldots, x_n^{s_n + t_n})) \\
        & = (x_1^{p(s_1 + t_1)}, x_2^{p(s_2 + t_2)}, \ldots, x_n^{p(s_n + t_n)}) \\
        & = (x_1^{ps_1} x_1^{pt_1}, x_2^{ps_2} x_2^{pt_2}, \ldots, x_n^{ps_n} x_n^{pt_n}) \\
        & = (x_1^{ps_1}, x_2^{ps_2}, \ldots, x_n^{ps_n})(x_1^{pt_1}, x_2^{pt_2}, \ldots, x_n^{pt_n}) \\
        & = \phi(x)\phi(y)
    \end{align*}
    as desired.
    \item $\phi$ collects all the $p$-powers, so $\im\phi = \langle x_1^p \rangle \times \langle x_2^p \rangle \times \cdots \times \langle x_n^p \rangle$. Moreover, $\ker\phi$ consists of all elements $(x_1^{s_1}, x_2^{s_2}, \ldots, x_n^{s_n})$ such that $x_1^{ps_1} = x_2^{ps_2} = \cdots = x_n^{ps_n} = e$, which implies that $ps_i$ is a multiple of $p^{\alpha_i}$ for all $i$, so $s_i$ is a multiple of $p^{\alpha_i - 1}$ for all $i$. Thus,
    \[\ker\phi = \left\langle x_1^{p^{\alpha_1-1}} \right\rangle \times \left\langle x_2^{p^{\alpha_2-1}} \right\rangle \times \cdots \times \left\langle x_n^{p^{\alpha_n-1}} \right\rangle.\]
    \item Note that $x_i^{p^{\alpha_i - 1}}$ has order $p$ for all $i$, so $\ker\phi$ is an elementary abelian group of order $p^n$. Moreover, note that each $\gen{x_i}$ is cyclic, hence every group is normal. Examining $\gen{x_i} / \gen{x_i^p}$, we see that $\gen{x_i} / \gen{x_i^p}$ is an elementary abelian group of order $p$. Since $A / \im\phi$ is the direct product of the $\gen{x_i} / \gen{x_i^p}$, we conclude that $A / \im\phi$ is an elementary abelian group of order $p^n$, by \hyperref[ex5.1.14]{Exercise 5.1.14}.
\end{solalph}

\begin{exercise}\label{ex5.2.8}
    Let $A$ be a finite abelian group (written multiplicatively) and let $p$ be a prime. Let
    \[A^p = \{a^p \mid a \in A\} \qquad \text{and} \qquad A_p = \{x \mid x^p = 1\}\]
    (so $A^p$ and $A_p$ are the image and kernel of the $p^{\text{th}}$-power map, respectively).
    \begin{subproblems}
        \item Prove that $A/A^p \cong A_p$. [Show that they are both elementary abelian and they have the same order.]
        \item Prove that the number of subgroups of $A$ of order $p$ equals the number of subgroups of $A$ of index $p$. [Reduce to the case where $A$ is an elementary abelian $p$-group.]
    \end{subproblems}
\end{exercise}

\begin{solalph}
    \item By Theorem 5.5, we may write $A$ as a direct product:
    \[A \cong A_{p_1} \times A_{p_2} \times \cdots \times A_{p_k},\]
    where each $A_{p_i}$ is a $p_i$-group for distinct primes $p_1, p_2, \ldots, p_k$. There are two cases:
    \begin{itemize}
        \item If $p \neq p_i$ for all $i$, then then $(p, p_i) = 1$ for all $i$, so the $p$-th power map is an automorphism of each $A_{p_i}$, so $A^p = A$ and $A_p = \{1\}$, so $A/A^p \cong A_p$.
        \item If $p = p_i$ for some $i$, then it is an automorphism for all $A_j$ for $j \neq i$, while the case for $A_{p_i}$ is handled by \hyperref[ex5.2.7]{Exercise 5.2.7}. Thus,
        \[A^p = A_{p_1} \times \cdots \times A_{p_{i-1}} \times A_{p_i}^p \times A_{p_{i+1}} \times \cdots \times A_{p_k}\]
        and
        \[A_p = \set 1 \times \cdots \times \set 1 \times (A_{p_i})_p \times \set 1 \times \cdots \times \set 1.\]
        We may again use \hyperref[ex5.1.14]{Exercise 5.1.14} along with the fact that $A_{p_i} / A_{p_i}^p \cong A_{p_i}^p$ to conclude that $A/A^p \cong A_p$.
    \end{itemize}
    \item Let $H \leq A$ such that $|A : H| = p$. Let $a \in A$. Then $aH \in A/H$. Since $|A : H| = p$, then $|aH| = 1$ or $p$. If $|aH| = 1$, then $a \in H$. If $|aH| = p$, then $a^pH = H$, or $a^p \in H$. In either case, we have $a^p \in H$, so $A^p \subseteq H$. By the Lattice Isomorphism Theorem, then subgroups of $A$ containing $A^p$ are in bijection with subgroups of $A/A^p$. By the previous part, $A/A^p \cong A_p$, so the number of index $p$ subgroups of $A$ equals the number of index $p$ subgroups of $A_p$.
    
    Now let $K \leq A$ where $|K| = p$. For any $k \in K$, we know $k^p = 1$, so $K \subseteq A_p$. Then subgroups of order $p$ in $A$ are exactly the subgroups of order $p$ in $A_p$, since $A_p \leq A$. Thus, the number of subgroups of order $p$ in $A$ equals the number of subgroups of order $p$ in $A_p$.

    Note that $A_p$ is an elementary abelian $p$-group, so it is isomorphic to $E_{p^n}$ for some $n$. By \hyperref[ex5.1.11]{Exercise 5.1.11}, the number of subgroups of order $p$ in $E_{p^n}$ is $\frac{p^n - 1}{p - 1}$. To see that the number of subgroups of index $p$ in $E_{p^n}$ is also $\frac{p^n - 1}{p - 1}$, we note that there are $p^n - 1$ surjective homomorphisms from $E_{p^n}$ to $E_p$, and each subgroup of index $p$ is the kernel of exactly $p - 1$ such homomorphisms, so there are $\frac{p^n - 1}{p - 1}$ subgroups of index $p$ in $E_{p^n}$.
\end{solalph}

\begin{exercise}\label{ex5.2.9}
    Let $A = Z_{60} \times Z_{45} \times Z_{12} \times Z_{36}$. Find the number of elements of order 2 and the number of subgroups of index 2 in $A$.
\end{exercise}

\begin{sol}
    Note that
    \begin{itemize}
        \item $Z_{60} \cong Z_4 \times Z_3 \times Z_5$,
        \item $Z_{45} \cong Z_9 \times Z_5$,
        \item $Z_{12} \cong Z_4 \times Z_3$, and
        \item $Z_{36} \cong Z_4 \times Z_9$.
    \end{itemize}
    Thus, the 2-power part of $A$ is $Z_4 \times Z_4 \times Z_4$. Elements of order 2 in $A$ are elements of the 2-power part as well as trivial parts of the other groups. The 2-power map sends $(a, b, c)$ to $(2a, 2b, 2c)$, so the kernel of the 2-power map is $\set{0, 2} \times \set{0, 2} \times \set{0, 2}$, which has 8 elements, 7 of which have order 2. Thus, there are 7 elements of order 2 in $A$.
\end{sol}

\begin{exercise}\label{ex5.2.10}
    Let $n$ and $k$ be positive integers and let $A$ be the free abelian group of rank $n$ (written additively). Prove that $A/kA$ is isomorphic to the direct product of $n$ copies of $\intmod[k]$ (here, $kA = \set{ka \mid a \in A}$). [See \hyperref[ex5.1.14]{Exercise 5.1.14}.]
\end{exercise}

\begin{sol}
    It is easy to see that $kA = k\z \times \cdots \times k\z$. Since $\z$ is abelian, then $k\z \nsub \z$. By \hyperref[ex5.1.14]{Exercise 5.1.14}, we have $A/kA \cong \intmod[k] \times \cdots \times \intmod[k].$
\end{sol}

\np

\begin{exercise}\label{ex5.2.11}
    Let $G$ be a nontrivial, finite abelian group of rank $t$.
    \begin{subproblems}
        \item Prove that the rank of $G$ equals the maximum of the ranks of its Sylow subgroups.
        \item Prove that $G$ can be generated by $t$ elements, but no subset with fewer than $t$ elements generates $G$. (One way of doing this is by using part(a) together with \hyperref[ex5.2.7]{Exercise 5.2.7}.)
    \end{subproblems}
\end{exercise}

\begin{solalph}
    \item Let $G$ be of type $(n_1, n_2, \ldots, n_t)$, and consider the invariant factor decomposition of $G$:
    \[G \cong Z_{n_1} \times Z_{n_2} \times \cdots \times Z_{n_t}.\]
    Let $p$ be a prime dividing the smallest factor $n_t$. Then the Sylow $p$-subgroup of $G$ is isomorphic to $Z_{p^{\alpha_1}} \times Z_{p^{\alpha_2}} \times \cdots \times Z_{p^{\alpha_t}}$, where $\alpha_i$ is the largest power of $p$ dividing $n_i$. Since $\alpha_t > 0$, the Sylow $p$-subgroup has rank $t$. On the other hand, if $q$ is any prime, then for any $Q \in \syl_q(G)$, we have
    \[Q \cong Z_{q^{\beta_1}} \times Z_{q^{\beta_2}} \times \cdots \times Z_{q^{\beta_t}},\]
    where $\beta_i \mid n_i$ for all $i$ with some potentially trivial factors, so the rank of $Q$ is at most $t$. Thus, the rank of $G$ equals the maximum of the ranks of its Sylow subgroups.
    \item It is clear that $G$ is generated by $t$ elements. Suppose $G = \gen{g_1, g_2, \ldots, g_{t-1}}$ and some $P \in \syl_p(G)$, which has rank $t$. Projecting $G$ onto $P$, we see that $P = \gen{\pi(g_1), \pi(g_2), \ldots, \pi(g_{t-1})}$. However, letting $\phi$ be the $p$-power map on $P$ as defined in \hyperref[ex5.2.7]{Exercise 5.2.7}, we have that $P/\im\phi \cong E_{p^t}$, so $P/\im\phi$ has rank $t$. Since $\pi(g_1), \pi(g_2), \ldots, \pi(g_{t-1})$ generate $P/\im\phi$, we have a contradiction. Thus, no subset with fewer than $t$ elements generates $G$.
\end{solalph}

\begin{specialexercise}\label{ex5.2.12}
    Let $n$ and $m$ be positive integers with $d = (n, m)$. Let $Z_n = \gen x$ and $Z_m = \gen y$. Let $A$ be the central product of $\gen x$ and $\gen y$ with an element of order $d$ identified, which has presentation
    \[\gen{x, y \mid x^n = y^m = 1, xy = yx, x^{n/d} = y^{m/d}}.\]
    Describe $A$ as a direct product of two cyclic groups.
    \down
    \noindent\textbf{Note}: While the problem specifically calls for two direct products, the proof shows that for any general $n, m$, $A$ is actually cyclic with order $Z_{nm/d}$, so the second direct product is trivial. I'm not sure if this is a typo, but just know that the problem will have a trivial direct product in the end.
\end{specialexercise}

\begin{sol}
    Since we have an element of order $d$ identified, it is easy to see that $|A| = nm/d$. By \hyperref[ex5.2.5]{Exercise 5.2.5}, we know that $\exp(Z_n \times Z_m) = \lcm(n, m) = nm/d$. Note that $|x| = n$ and $|y| = m$ in $A$ still, for otherwise there would be a contradiction to the order of $A$ as we would not be able to account for all the elements of $A$.

    Again, by \hyperref[ex5.2.5]{Exercise 5.2.5}, we know that $A$ contains an element of order $n$ iff $n \mid \exp(A)$, and $A$ contains an element of order $m$ iff $m \mid \exp(A)$. Since $\exp(A) \mid |A| = nm/d$, we have that $\exp(A)$ is a common multiple of $n$ and $m$, so $\exp(A)$ is a multiple of $\lcm(n, m) = nm/d$, hence $\lcm(n, m) \mid \exp(A)$. On the other hand, $\exp(A) \mid |A| = nm/d$ since the order of any element divides the order of the group, so $\exp(A) \mid \lcm(n, m)$. Thus, $\exp(A) = \lcm(n, m)$. We now conclude that $A$ is cyclic, so $A \cong Z_{nm/d} \times Z_1$.
\end{sol}

\np

\begin{exercise}\label{ex5.2.13}
    Let $A = \gen{x_1} \times \gen{x_2} \times \cdots \times \gen{x_r}$ be a finite abelian group with $|x_i| = n_i$ for $1 \leq i \leq r$. Find a presentation for $A$. Prove that if $G$ is any group containing commuting elements $g_1, \ldots, g_r$ such that $g_i^{n_i} = 1$ for $1 \leq i \leq r$, then there is a unique homomorphism from $A$ to $G$ which sends $x_i$ to $g_i$ for all $i$.
\end{exercise}

\begin{sol}
    A presentation for $A$ is given by
    \[\gen{x_1, x_2, \ldots, x_r \mid x_i^{n_i} = 1 \text{ for } 1 \leq i \leq r, x_ix_j = x_jx_i \text{ for } 1 \leq i < j \leq r}.\]
    Now, observe that every element of $A$ is of the form $x_1^{s_1} x_2^{s_2} \cdots x_r^{s_r}$ for some integers $s_1, s_2, \ldots, s_r$. Define $\phi : A \to G$ by $\phi(x_1^{s_1} x_2^{s_2} \cdots x_r^{s_r}) = g_1^{s_1} g_2^{s_2} \cdots g_r^{s_r}$. Since the $g_i$ commute, $\phi$ is well-defined. Moreover, $\phi$ is a homomorphism since
    \begin{align*}
        \phi((x_1^{s_1} x_2^{s_2} \cdots x_r^{s_r})(x_1^{t_1} x_2^{t_2} \cdots x_r^{t_r})) & = \phi(x_1^{s_1 + t_1} x_2^{s_2 + t_2} \cdots x_r^{s_r + t_r}) \\
        & = g_1^{s_1 + t_1} g_2^{s_2 + t_2} \cdots g_r^{s_r + t_r} \\
        & = (g_1^{s_1} g_2^{s_2} \cdots g_r^{s_r})(g_1^{t_1} g_2^{t_2} \cdots g_r^{t_r}) \\
        & = \phi(x_1^{s_1} x_2^{s_2} \cdots x_r^{s_r})\phi(x_1^{t_1} x_2^{t_2} \cdots x_r^{t_r})
    \end{align*}
    for all integers $s_1, s_2, \ldots, s_r$ and $t_1, t_2, \ldots, t_r$. Finally, $\phi(x_i) = g_i$ for all $i$ by definition. To see that $\phi$ is the unique homomorphism from $A$ to $G$ sending $x_i$ to $g_i$ for all $i$, let $\psi : A \to G$ be any such homomorphism. Then $\psi(x_1^{s_1} x_2^{s_2} \cdots x_r^{s_r}) = \psi(x_1)^{s_1} \psi(x_2)^{s_2} \cdots \psi(x_r)^{s_r} = g_1^{s_1} g_2^{s_2} \cdots g_r^{s_r}$ for all integers $s_1, s_2, \ldots, s_r$, so $\psi = \phi$.
\end{sol}

\begin{specialexercise}\label{ex5.2.14}
    For any group $G$ define the \textit{dual group} of $G$ (denoted $\wh{G}$) to be the set of all homomorphisms from $G$ into the multiplicative group of roots of unity in $\mathbb{C}$. Define a group operation in $\wh{G}$ by pointwise multiplication of functions: if $\chi, \psi$ are homomorphisms from $G$ into the group of roots of unity then $\chi\psi$ is the homomorphism given by $(\chi\psi)(g) = \chi(g)\psi(g)$ for all $g \in G$, where the latter multiplication takes place in $\mathbb{C}$.
    \begin{subproblems}
        \item Show that this operation on $\wh{G}$ makes $\wh{G}$ into an abelian group. [Show that the identity is the map $g \mapsto 1$ for all $g \in G$ and the inverse of $\chi \in \wh{G}$ is the map $g \mapsto \chi(g)^{-1}$.]
        \item If $G$ is a finite abelian group, prove that $\wh{G} \cong G$. [Write $G$ as $\langle x_1 \rangle \times \cdots \times \langle x_r \rangle$ and if $n_i = |x_i|$ define $\chi_i$ to be the homomorphism which sends $x_i$ to $e^{2\pi i/n_i}$ and sends $x_j$ to $1$, for all $j \neq i$. Prove $\chi_i$ has order $n_i$ in $\wh{G}$ and $\wh{G} = \langle \chi_1 \rangle \times \cdots \times \langle \chi_r \rangle$.]
    \end{subproblems}
\end{specialexercise}

\begin{solalph}
    \item Multiplication in $\c$ is already associative and commutative, so pointwise multiplication of functions is also associative and commutative. The map $\ep : G \to \c$ given by $\ep(g) = 1$ for all $g \in G$ is a homomorphism, since $\ep(gh) = 1 = 1 \cdot 1 = \ep(g)\ep(h)$ for all $g, h \in G$, and 1 is a root of unity. Moreover, for any $\chi \in \wh G$, then $(\chi\ep)(g) = \chi(g)\ep(g) = \chi(g) \cdot 1 = \chi(g)$ for all $g \in G$. Thus, $\ep$ is the identity element of $\wh{G}$. Finally, if $\chi \in \wh{G}$, then the map $\chi^{-1} : G \to \c$ given by $\chi^{-1}(g) = \chi(g)^{-1}$ is a homomorphism, since $\chi^{-1}(gh) = \chi(gh)^{-1} = (\chi(g)\chi(h))^{-1} = \chi(g)^{-1}\chi(h)^{-1} = \chi^{-1}(g)\chi^{-1}(h)$ for all $g, h \in G$, and $\chi(g)^{-1}$ is a root of unity since $\chi(g)$ is a root of unity. Moreover, $\chi\chi^{-1} = \ep$, so $\chi^{-1}$ is the inverse of $\chi$ in $\wh{G}$.
    \item To see that $\chi_i$ has order $n_i$ in $\wh G$, let $g = x_1^{s_1} \cdots x_r^{s_r} \in G$. Then
    \[\chi_i(g) = \chi_i(x_1^{s_1} \cdots x_r^{s_r}) = \chi_i(x_1)^{s_1} \cdots \chi_i(x_r)^{s_r} = e^{2\pi i s_i / n_i},\]
    where all terms $\chi_i(x_j)^{s_j}$ for $j \neq i$ are 1. Thus, $\chi_i^k(g) = e^{2\pi i k s_i / n_i}$ for all integers $k$, so $\chi_i^k = \ep$ if and only if $n_i \mid k$. Thus, $\chi_i$ has order $n_i$ in $\wh{G}$. Now, note by \hyperref[ex5.2.13]{Exercise 5.2.13} that there exists a unique homomorphism $\phi : G \to \wh G$ such that $\phi(x_i) = \chi_i$ for all $i$. We now show that $\phi$ is an isomorphism.

    Let $g = x_1^{s_1} \cdots x_r^{s_r} \in G$ such that $\phi(g) = \ep$. Then for any $x_i$, we have that
    \[\phi(g)(x_i) = (\chi_1^{s_1} \cdots \chi_r^{s_r})(x_i) = \chi_i^{s_i}(x_i) = e^{2\pi i s_i / n_i} = 1.\]
    Thus, $n_i \mid s_i$ for all $i$, so $g = x_1^{s_1} \cdots x_r^{s_r} = 1$. Therefore, $\phi$ is injective. Similarly, for any $\chi \in \wh G$, and $x_i^{n_i} = 1$ for some $i$, then $\chi(x_i)^{n_i} = \chi(x_i^{n_i}) = \chi(1) = 1$, so $\chi(x_i)$ is an $n_i$-th root of unity, so $\chi(x_i) = e^{2\pi i s_i / n_i}$ for some integer $s_i$. Thus, $\chi = \phi(x_1^{s_1} \cdots x_r^{s_r})$, so $\phi$ is surjective. Therefore, $\phi$ is an isomorphism, so $\wh{G} \cong G$.
\end{solalph}

\begin{exercise}\label{ex5.2.15}
    Let $G = \gen x \times \gen y$, where $|x| = 8$ and $|y| = 4$.
    \begin{subproblems}
        \item Find all pairs $a, b$ in $G$ such that $G = \gen a \times \gen b$ (where $a$ and $b$ are expressed in terms of $x$ and $y$).
        \item Let $H = \gen{x^2y, y^2} \cong Z_4 \times Z_2$. Prove that there are no elements $a, b$ of $G$ such that $G = \gen a \times \gen b$ and $H = \gen{a^2} \times \gen{b^2}$ (i.e., one cannot pick direct product generators for $G$ in such a way that some powers of these are direct product generators for $H$).
    \end{subproblems}
\end{exercise}

\begin{solalph}
    \item From the description of $G$, we see that $G$ is of type $(8, 4)$, so we have the following conditions for $G = \gen a \times \gen b$:
    \begin{itemize}
        \item $|a| = 8$ and $|b| = 4$,
        \item $\gen a \cap \gen b = 1$.
    \end{itemize}
    Let $a = x^s y^t$ and $b = x^u y^v$ for some integers $s, t, u, v$. Then $|a| = \lcm(8/\gcd(8, s), 4/\gcd(4, t))$ and $|b| = \lcm(8/\gcd(8, u), 4/\gcd(4, v))$. For $|a|$ to be 8, we must have $\gcd(8, s) = 1$, implying that $s$ must be odd, while $t$ may range freely. For $|b|$ to be 4, then $8/\gcd(8, u) \mid 4$ and $\max(8/\gcd(8, u), 4/\gcd(4, v)) = 4$, so we have the following cases:
    \begin{enumerate}[label=(\arabic*)]
        \item If $\gcd(8, u) = 8$, then $u = 0$ and $v$ must be odd.
        \item If $\gcd(8, u) = 4$, then $u = 4$ and $v$ must be odd.
        \item If $\gcd(8, u) = 2$, then $u = 2, 6$ and $v$ can be free.
    \end{enumerate}
    To have trivial intersection, we must have that $x^{ks} y^{kt} = x^{\ell u} y^{\ell v} = 1$, or $ks \equiv \ell u \bmod 8$ and $kt \equiv \ell v \bmod 4$ implies that $k \equiv 0 \bmod 8$ and $\ell \equiv 0 \bmod 4$, which is seen from $a^k = 1$ and $b^\ell = 1$ iff $8 \mid k$ and $4 \mid \ell$. Thus, we check the cases for $|b|$:
    \begin{enumerate}[label=(\arabic*)]
        \item $u = 0$ and $v$ is odd: We have $ks \equiv 0 \bmod 8$ and $kt \equiv \ell v \bmod 4$. Since $s$ is odd, then $k \equiv 0 \bmod 8$. Then $kt \equiv 0 \bmod 4$, hence $\ell v \equiv 0 \bmod 4$. Since $v$ is odd, then $\ell \equiv 0 \bmod 4$, so we have trivial intersection.
        \item $u = 4$ and $v$ is odd: We have $ks \equiv 4\ell \bmod 8$ and $kt \equiv \ell v \bmod 4$. Since $s$ is odd, then $k \equiv 4\ell s\inv \bmod 8$. Note that $k \equiv 0 \bmod 4$, so $kt \equiv 0 \bmod 4$, hence $\ell v \equiv 0 \bmod 4$. Since $v$ is odd, then $\ell \equiv 0 \bmod 4$, so we have trivial intersection.
        \item $u = 2$ and $v$ free (the case $u = 6$ and $v$ free is similar): We have $ks \equiv 2\ell \bmod 8$ and $kt \equiv \ell v \bmod 4$. Since $s$ is odd, then $k \equiv 2 \ell s\inv \bmod 8$. We know that intersection will be nontrivial when $k \not\equiv 0 \bmod 8$ and $\ell \not\equiv 0 \bmod 4$. By inspection, we see that $\ell = 2$ gives $k \equiv 4s\inv \equiv 4 \bmod 8$ and $4t \equiv 2v \bmod 4$, which has nontrivial solution when $v$ is even. Inspecting odd $v$, we see that $k \equiv 2\ell s\inv \bmod 8$ reduced mod 4 gives $k \equiv 2 \ell s\inv \bmod 4$. Then $2 \ell s\inv \equiv \ell v \bmod 4$ implies that $\ell(2s\inv - v) \equiv 0 \bmod 4$. Since $v$ is odd, then $2s\inv - v$ is odd, so $\ell \equiv 0 \bmod 4$, so we have trivial intersection.
    \end{enumerate}
    Therefore, the pairs $a, b$ such that $G = \gen a \times \gen b$ are given by
    \begin{itemize}
        \item $a = x^s y^t$ and $b = y^v$ for some odd integers $s, v$ and any integer $t$,
        \item $a = x^s y^t$ and $b = x^4 y^v$ for some odd integers $s, v$ and any integer $t$,
        \item $a = x^s y^t$ and $b = x^2 y^v$ for some odd integer $s$, any integer $t$, and any odd integer $v$, or
        \item $a = x^s y^t$ and $b = x^6 y^v$ for some odd integer $s$, any integer $t$, and any odd integer $v$.
    \end{itemize}
    \item We begin with the elements of $H$:
    \[H = \set{1, x^2y, x^4y^2, x^6y^3, y^2, x^2y^3, x^4y, x^6y^2}\]
    Note that $H = \gen{x^2y} \times \gen{y^2}$, so it suffices to check whether or not $a^2$ generates $\gen{x^2y}$ from the choices of $a$ from part (a). Note that $a^2 = x^{2s} y^{2t}$, which implies that $y$ will always have an even exponent. However, the only element with even exponent on $y$ from $\gen{x^2y}$ is with the term $x^4y^2$, which would imply that $\gen{x^4y^2} = \gen{x^2y}$, which is a contradiction since $x^2y$ has order 4 while $x^4y^2$ has order 2. Thus, there are no elements $a, b$ of $G$ such that $G = \gen a \times \gen b$ and $H = \gen{a^2} \times \gen{b^2}$.
\end{solalph}

\np

\begin{exercise}\label{ex5.2.16}
    Prove that no finitely generated abelian group is divisible (cf. \hyperref[ex2.4.19]{Exercise 2.4.19}).
\end{exercise}

\begin{sol}
    Let $A$ be a finitely generated abelian group. By Theorem 5.5, we can write $A$ as a direct product of cyclic groups:
    \[A \cong \z^r \times Z_{n_1} \times Z_{n_2} \times \cdots \times Z_{n_t},\]
    where $n_i > 1$ for all $i$ and $r \geq 0$. For $r > 0$, observe that $\z$ is not divisible, since there is no $x \in \z$ such that $2x = 1$. For $r = 0$, observe that $A$ is a finite abelian group. By \hyperref[ex2.4.19]{Exercise 2.4.19}, finite abelian groups are not divisible. Thus, no finitely generated abelian group is divisible.
\end{sol}

\newpage

\subsection{Table of Groups of Small Order}

\begin{exercise}\label{ex5.3.1}
    Prove that $D_{16}, Z_2 \times D_8, Z_2 \times Q_8, Z_4 * D_8, QD_{16}$, and $M$ are nonisomorphic, non-abelian groups of order 16 (where $Z_4 * D_8$ is described in \hyperref[ex5.1.12]{Exercise 5.1.12} and $QD_{16}$ and $M$ are described in \hyperref[ex2.5.11]{Exercise 2.5.11} and \hyperref[ex2.5.14]{Exercise 2.5.14} respectively).
\end{exercise}

\begin{sol}
    We do a ``decision tree''-like process to show that these are non-isomorphic:
    \begin{itemize}
        \item $D_{16}$
        \begin{itemize}
            \item $Z_2 \times D_8$ has no element of order 8, while $D_{16}$ does.
            \item $Z_2 \times Q_8$ has no element of order 8, while $D_{16}$ does.
            \item $Z_4 * D_8$ has no elements of order 8, since quotienting preserves only divisibility of orders from $Z_4 \times D_8$, which has no elements of order 8, while $D_{16}$ does.
            \item In $QD_{16}$, we have $|\sigma^4| = 2$, and $|\sigma^k\tau| = 2$ when
            \[\sigma^k\tau\sigma^k\tau = \sigma^k\sigma^{3k}\tau\tau = \sigma^{4k},\]
            hence $\sigma^{4k} = 1$ when $k$ is even, hence there are only 5 elements of order 2 in $QD_{16}$, while there are 8 reflections of order 2, plus the central rotation $r^4$, for a total of 9 elements of order 2 in $D_{16}$.
            \item $M$ has $|u| = 2$, with
            \[uv^kuv^k = uuv^{5k}v^k = v^{6k},\]
            so $v^{6k} = 1$ when $8 \mid 6k$, or $4 \mid 3k$, or $4 \mid k$. Then only $k = 4$ works, giving $M$ a total of 3 elements of order 2.
        \end{itemize}
        \item $Z_2 \times D_8$
        \begin{itemize}
            \item $Z_2 \times Q_8$ does not have the same number of elements of order 2 as $Z_2 \times D_8$ does, since $D_8$ has 5 elements of order 2, while $Q_8$ has only 1 element of order 2.
            \item Note that $\bar{(x^k, q)} \in Z_4 * Q_8 \cong Z_4 * D_8$ has order 2 when $(x^k, q)^2 = (1, 1)$ or $(x^2, -1)$. Since we already identify $x^2$ with $-1$ in the quotient, then the case of $q^2 = 1$ and $2i \equiv 0 \bmod 4$ gives the same element, hence only 1 element of order 2. The other case of $q^2 = -1$ and $2i \equiv 2 \bmod 4$, or $i$ odd and $q = Q_8 \setminus Z(Q_8)$ gives 6 elements of order 2, for a total of 7 elements of order 2 in $Z_4 * D_8$, while $Z_2 \times D_8$ has 11 elements of order 2.
            \item $QD_{16}$ contains elements of order 8, while $Z_2 \times D_8$ does not.
            \item $M$ contains elements of order 8, while $Z_2 \times D_8$ does not.
        \end{itemize}
        \item $Z_2 \times Q_8$
        \begin{itemize}
            \item Note that $Z_2 \times Q_8$ only has 3 elements of order 2, while $Z_4 * D_8$ has 7 elements of order 2.
            \item $QD_{16}$ contains elements of order 8, while $Z_2 \times Q_8$ does not.
            \item $M$ contains elements of order 8, while $Z_2 \times Q_8$ does not.
        \end{itemize}
        \item $Z_4 * D_8$
        \begin{itemize}
            \item $QD_{16}$ contains elements of order 8, while $Z_4 * D_8$ does not.
            \item $M$ contains elements of order 8, while $Z_4 * D_8$ does not.
        \end{itemize}
        \item $QD_{16}$
        \begin{itemize}
            \item $M$ has only 3 elements of order 2, while $QD_{16}$ has 5 elements of order 2. \qh
        \end{itemize}
    \end{itemize}
\end{sol}